On centrality of even orthogonal $\Kt$

### 2016-05-24 paper.tex
Purged commented lines

%\begin{thebibliography}{99}
%\bibitem{Lav2} A.~Lavrenov, A local--global principle for symplectic $\mathrm K_2$, {\it to appear}.
%\bibitem{Rao} *Statement of Ravi Rao's hypothesis
%\bibitem{Swa1} *Swan's relative Steinberg group 1
%\bibitem{Swa2} *Swan's relative Steinberg group 2
%\bibitem{Swa3} *Swan's relative Steinberg group 3
%\bibitem{Tul0} *Tulenbaev's paper on almost commutative rings (cite somewhere)
%\end{thebibliography}

@@

The paper is organised as follows. 
After introducing the preliminaries we explain in \cref{sec:patching} how the patching technique of~\cite{SCh} allows one to reduce \cref{tm:lg-principle} to the special case $\Phi=\rA_3$.
Next, in \cref{sec:yap} we describe a new ``transpose-symmetric'' presentation for the relative linear Steinberg group.
Finally, this presentation is used in~\cref{sec:lgp} to prove \cref{tm:lg-principle} in the case $\Phi=\rA_3$.
%It mimics the standart vector representation of $\SL(n,\,R)$, therefore in the first section we use representations and corresponding vector notation. 
%Finally, in \cref{sec:patching} we derive our principal results from this special case and the amalgamation theorem for relative Steinberg groups from~\cite{SCh}.
%There we switch to the root notation.

@@

%Tulenbaev used {\it another presentation} for a relative Steinberg group to prove a local--global
%principle for a linear Steinberg group~\cite{Tul}. 
%This relative another presentation was inspired by the {\it absolute} one of van der Kallen~\cite{vdK}. 
%However, Tulenbaev's presentation can only be used for $\St(l+1,\,R)=\St(\mathrm A_l,\,R)$, $l\geq4$.

%Tulenbaev's result can be applied to the subsystems of type $\mathrm A_4$ inside $\mathrm E_l$ and ``glued'' 
% to get a local--global principle for Steinberg groups of type $\mathrm E_l$~\cite{SCh}. 
%To use this glueing method for groups of type $\mathrm D_l$ as well one has to work with subsystems 
%of type $\mathrm A_3$ instead, and thus extend Tulenbaev's result to the case of $\mathrm A_3$.

%Recently, the local--global principle was obtained for Steinberg groups of type $\mathrm C_l$ starting from rank 3~\cite{Lav2}. 
%That is another motivation to consider the linear Steinberg group in rank 3.

@@

In \cite{Sus} Suslin showed that for $n\geq 3$ the elementary group $\E(n, R)$ coincides with the subgroup of $\GL(n,\,R)$ generated by matrices
of the form $t(u,\,v)=1+uv^t$ where $u\in\Um(n,\,R)$, $v\in R^n$ and $u^tv=0$. Here $u^t$ stands for the transpose of $u$.
This result clearly implies that $\E(n,\,R)$ is normal in $\GL(n,\,R)$.
Developing Suslin's ideas, W.~van~der~Kallen showed that for $n\geq4$ the Steinberg group $\St(n,\,R)$ is isomorphic to the group with the following presentation by generators
$$\{X(u,\,v)\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\}$$ and relations:
\setcounter{equation}{0} \renewcommand{\theequation}{K\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add1} \\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'). \label{conj1}
\end{align}
This presentation clearly implies that $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ is a central extension.
%Our notation for the generators of $\St(n, R)$ differs from van der Kallen's, e.\,g.
Notice that we parametrize elements $X(u, v)$ by two columns rather than by a column and a row as it is done in~\cite{vdK}.
Our generator $X(u,v)$ corresponds to van der Kallen's generator $X(u, v^t)$.

@@

%This $T$ is an inverse to a map induced by localisation, so that constructing $\mathrm T$ may be treated
%as a proof of excision property for Steinberg group in this very particular situation.

%To prove the local--global principle for Steinberg groups of type $\mathrm D_l$ and centrality 
%of $\mathrm K_2(\mathrm D_l,\,R)$ we need such a map for $n=4$ (i.e., for the group of type $\mathrm A_3$).

%Since there are known counterexamples to centrality of $\Kt$ and, thus, to \cref{tm:lg-principle} for $\Phi$ of rank $2$ (see~\cite{W}), 
%this gives a complete answer to the question for which $\ell$ the local-global principle holds for $\Kt(\rC_\ell, R)$.

%The main idea of our proof of~\cref{tm:lg-principle} is to first settle the case $\Phi=\rA_3$ and then derive the other cases from it using the patching technique of~\cite{SCh}.

%Compare \cref{tm:lg-principle} with the local-global principle for $\Ko$ which states that an element of the congruence subgroup $g\in \GG(\Phi,\,R[X], XR[X])$ is elementary
%(i.\,e. lies in $\E(\Phi, R[X])$) if and only if $\lambda_M^*(g)=1\in\E(\Phi,\,R_M[X])$ for all maximal ideals $M \trianglelefteq R$.

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
\setcounter{lm}{1}
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ and $n\geq 4$ the group $\St(n,\,R,\,I)$ is isomorphic to the group defined by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ and the following set of relations:
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
Moreover, the group $\St(n, R)$ decomposes into the semidirect product $\St(n,\,R, I)\rtimes \St(n,\,R/I)$.
\end{prop}
Here, $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.
In the definition of the semidirect product the group $\St(n, R/I)$ acts on $\St(n, R, I)$ by conjugation via the splitting map, i.\,e. ${}^g X(u,\,v) = X(\phi(\sigma(g))u, \phi(\sigma(g))^*v)$ for $g \in \St(n, R/I)$.

@@

Both ``another presentation'' of van der Kallen and that of Tulenbaev are given in terms of generators parametrised by pairs of vectors,
 where the first one is ``nice'' in some sense i.\,e. unimodular, a column of an elementary matrix or the like, while the second one is arbitrary.
It is easy to formulate the additivity property in the second argument for these generators, cf.~\eqref{add1}, \eqref{add2},
while it is not so easy when it comes to the additivity in the first argument, cf. \eqref{add3}.
%Such presentations allowed them to formulate easily an additivity by the second argument, but one has to write something similar to T3 instead of additivity by the first one.

@@

\begin{prop} \label{lm:map-iota}
 For a splitting ideal $I \trianglelefteq R$ and $n\geq 4$ the groups $\St^*(n,\,R,\,I)$ and $\St(n,\,R, I)$ are isomorphic.
% there exists a well-defined homomorphism $$\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R, I) \subseteq \St(n, R)$$ 
% satisfying the equalities $\iota (F(u,\,v)) = X(u,\,v)$ and $\iota (S(u,\,v)) = Y(u,\,v)$. 
\end{prop}
%\begin{proof}
%We already know that $\iota$ is well-defined. It suffices to check that the image of $\iota$ is contained in $\St(n,\,R,I)$.
%Denote the canonical projection $R\epi R/I$ by $\rho$.
%Clearly, the induced group morphism $\rho^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$ sends $X(u,\,v)$ to $X(\rho(u),\,\rho(v))$ and $Y(u,\,v)$ to $Y(\rho(u),\,\rho(v))$. 
%Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective, therefore for $x\in\E(n,\,R)e_1$ one has $\rho(x)\in\E(n,\,R/I)e_1$. 
%, as required. \end{proof}

\begin{comment}
We constructed $\iota$ for an arbitrary $I\trianglelefteq R$. For $I$ a splitting ideal we construct an inverse homomorphism
$$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I).
$$
With this end we recall a presentation of $\St(n,\,R,\,I)$ as a group with action of $\St(n,\,R)$. For a splitting ideal such a presentation was firstly obtained by Swan~\cite{Swa1,Swa2,Swa3}, and for an arbitrary $I$ by Keune~\cite{Keu} and Loday~\cite{Lod}.

For a group $G$ acting on a group $H$ from the left denote by $\!\,^gh$ the image of $h\in H$ under homomorphism corresponding by $g\in G$.

Let $I\trianglelefteq R$ be a splitting ideal then $\St(n,\,R,\,I)$ can be presented as a group with action of $\St(n,\,R)$ with the set of (relative) generators
$$
\{y_{ij}(a)\mid a\in I,\ 1\leq i\neq j\leq n\}
$$
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{KL\arabic{equation}}
\begin{align}
&y_{ij}(a)y_{ij}(b)=y_{ij}(a+b),\\
&\!\,^{x_{ij}(r)}y_{hk}(b)=y_{hk}(b),\text{ for }h\neq j,\ k\neq i,\\
&\!\,^{x_{ij}(r)}y_{jk}(b)\cdot y_{jk}(-b)=y_{ik}(rb),\\
&y_{ij}(a)\cdot\,^{x_{jk}(s)}y_{ij}(-a)=y_{ik}(as),\\
&\!\,^{x_{hk}(a)}\big(\!\,^gy_{ij}(b)\big)=y_{hk}(a)\cdot\,^gy_{ij}(b)\cdot y_{hk}(-a)\,\text{ for }a\in I,\ g\in\St(n,\,R).
\end{align}

In other words, $\St(n,\,R,\,I)$ is isomorphic to a quotient group of a free group $F$ on the set of generators 
$$\St(n,\,R)\times\{y_{ij}(a)\mid a\in I,\ 1\leq i\neq j\leq n\}$$
(the natural action of $\St(n,\,R)$ on $F$ is given by $\!\,^f(g,\,y)=(fg,\,y)$) modulo the normal equivariant subgroup generated by KL1--KL5 (where $y_{ij}(a)$ stands for $(1,\,y_{ij}(a))$. Keune in Loday consider only stable Steinberg group $\St(R)=\varinjlim\St(n,\,R)$. This presentation for unstable group firstly appeared in~\cite{SCh} in the context of Chevalley groups. Observe that the definition of the relative Steinberg group in~\cite[Def.~3.3]{SCh} differs from ours but remark after this definition and~\cite[Lem.~8]{SCh} immediantly imply that for a splitting ideal $I\trianglelefteq R$ it coincides with the one given in the present paper. The existance of the presentation is proven in~\cite[Prop.~6]{SCh}. Relation~5 in~\cite[Prop.~6]{SCh} joins relations KL3 and KL4, and relations~2 and 3 follow from KL4 and KL5 with the use of KL6.

One could probably wonder why Tulenbaev needs his own presentation for $\St(n,\,R,\,I)$ and does not use the one of Keune--Loday. The reason is the following. To construct a map from Keune--Loday group one should firstly define an action of the absolute Steinberg group on the target group, in our context, the action of $\St(n,\,B_a)$ on $\St(n,\,B,\,I)$, what is probably possible, but seems to be much harder, then giving another presentation (in particular, one should define $\!\,^{x_{ij}(r/a^m)}y_{ji}(c)$).

Now, to construct
$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I)
$
we need to define an action of $\St(n,\,R)$ on $\St^*(n,\,R,\,I)$ and find elements $y^*_{ij}(a)\in\St^*(n,\,R,\,I)$ subject to KL1--KL5. To define a action of the absolute group we use van der Kallen's another presentaion for it. For $u\in\Um(n,\,R)$, $v\in R^n$, $u^tv=0$ define
$
\alpha(u,\,v)\colon\St^*(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I)
$
by $\alpha(u,\,v)\big(F(u',\,v')\big)=F(t(u,\,v)u',\,t(u,\,v)^*v')$, and $\alpha(u,\,v)\big(S(u',\,v')\big)=S(t(u,\,v)u',\,t(u,\,v)^*v')$. Obviously, the images of the generators satisfy R1--R4, so that $\alpha(u,\,v)$ is a well-defined automorphism. Also, $\alpha$'s themselves clearly satisfy K1--K2, thus $X(u,\,v)\mapsto\alpha(u,\,v)$ is a well-defined action of $\St(n,\,R)$ on $\St^*(n,\,R,\,I)$.

Next, define $y_{ij}^*(a)=F(e_i,\,e_ja)$, $a\in I$. These elements obviously satisfy KL1, KL2 and KL5. Check KL3:
\begin{multline*}
\!\,^{x_{ij}(r)}y^*_{jk}(b)=F(t_{ij}(r)e_j,\,t_{ji}(-r)e_kb)=F(e_ir+e_j,\,e_kb)=\\
=S(e_irb+e_jb,\,e_k)=S(e_irb,\,e_k)S(e_jb,\,e_k)=F(e_i,\,e_krb)F(e_j,\,e_kb).
\end{multline*} 
KL4 is similar. Finally, we have a well-defined map
$$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I)
$$
sending $y_{ij}(a)$ to $y^*_{ij}(a)$. 
\end{comment}

\section{The local-global principle for \texorpdfstring{$\Kt$}{K2}}\label{sec:lgp}
%At this point we have demonstrated for $n\geq4$ and a splitting ideal $I$ that $\St^*(n,\,R,\,I)$ and $\St(n,\,R,\,I)$ are isomorphic.
Now we turn to the main result of this paper, namely, for $n\geq 4$ we construct the map $$T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X]),$$ that fits into the diagram \eqref{eq:diagram}.

@@

For a vector $u\in R^n$ we denote by $I(u)$ the ideal generated by the components of $u\in R^n$, i.\,e. $I(u)=\sum\limits_{k=1}^nu_kR$.
To prove \cref{a3map} we follow the approach of Tulenbaev and construct yet another family of elements $X_{u,v}(a)\in \St(n,\,B)$, $a\in B$ which contains van der Kallen's generators $X(u, v)$ as a subfamily, i.\,e. $X_{u,v}(1) = X(u, v)$.
The key feature of this definition is that the assumption $u \in\Um(n, B)$ is replaced by a weaker condition $a \in I(u)$ equivalent to imposing that $u$ becomes unimodular after localisation by the powers of $a$.
%The idea is the following. Say $a$ is not a zero divisor, then $ua^m\in B$ for some $m$ and we will construct some element ``$X(ua^m,\,va^{-m})$''.
%One can not garantee that $ua^m\in\E(n,\,B)e_1a^k$, however, the ideal generated by entries of $ua^m$ contains some power of $a$ .
%More formally, we have the following definition.
%Then, we need generators parametrised by pairs $(u,\,v)$ with $u^tv=0$, $v\in I$, $a^m\in\mathrm I(u)$ for some $m\in\mathbb N$. Such generators are defined in~\cite{Tul}.

>>>>

@@

The paper is organised as follows. 
After introducing the preliminaries we explain in \cref{sec:patching} how the patching technique of~\cite{SCh} allows one to reduce \cref{tm:lg-principle} to the special case $\Phi=\rA_3$.
Next, in \cref{sec:yap} we describe a new ``transpose-symmetric'' presentation for the relative linear Steinberg group.
Finally, this presentation is used in~\cref{sec:lgp} to prove \cref{tm:lg-principle} in the case $\Phi=\rA_3$.

@@

@@

In \cite{Sus} Suslin showed that for $n\geq 3$ the elementary group $\E(n, R)$ coincides with the subgroup of $\GL(n,\,R)$ generated by matrices
of the form $t(u,\,v)=1+uv^t$ where $u\in\Um(n,\,R)$, $v\in R^n$ and $u^tv=0$. Here $u^t$ stands for the transpose of $u$.
This result clearly implies that $\E(n,\,R)$ is normal in $\GL(n,\,R)$.
Developing Suslin's ideas, W.~van~der~Kallen showed that for $n\geq4$ the Steinberg group $\St(n,\,R)$ is isomorphic to the group with the following presentation by generators
$$\{X(u,\,v)\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\}$$ and relations:
\setcounter{equation}{0} \renewcommand{\theequation}{K\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add1} \\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'). \label{conj1}
\end{align}
This presentation clearly implies that $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ is a central extension.
Notice that we parametrize elements $X(u, v)$ by two columns rather than by a column and a row as it is done in~\cite{vdK}.
Our generator $X(u,v)$ corresponds to van der Kallen's generator $X(u, v^t)$.

@@

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
\setcounter{lm}{1}
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ and $n\geq 4$ the group $\St(n,\,R,\,I)$ is isomorphic to the group defined by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ and the following set of relations:
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
Moreover, the group $\St(n, R)$ decomposes into the semidirect product $\St(n,\,R, I)\rtimes \St(n,\,R/I)$.
\end{prop}
Here, $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.
In the definition of the semidirect product the group $\St(n, R/I)$ acts on $\St(n, R, I)$ by conjugation via the splitting map, i.\,e. ${}^g X(u,\,v) = X(\phi(\sigma(g))u, \phi(\sigma(g))^*v)$ for $g \in \St(n, R/I)$.

@@

Both ``another presentation'' of van der Kallen and that of Tulenbaev are given in terms of generators parametrised by pairs of vectors,
 where the first one is ``nice'' in some sense i.\,e. unimodular, a column of an elementary matrix or the like, while the second one is arbitrary.
It is easy to formulate the additivity property in the second argument for these generators, cf.~\eqref{add1}, \eqref{add2},
while it is not so easy when it comes to the additivity in the first argument, cf. \eqref{add3}.

@@

\begin{prop} \label{lm:map-iota}
 For a splitting ideal $I \trianglelefteq R$ and $n\geq 4$ the groups $\St^*(n,\,R,\,I)$ and $\St(n,\,R, I)$ are isomorphic.
\end{prop}

\section{The local-global principle for \texorpdfstring{$\Kt$}{K2}}\label{sec:lgp}
Now we turn to the main result of this paper, namely, for $n\geq 4$ we construct the map $$T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X]),$$ that fits into the diagram \eqref{eq:diagram}.

@@

For a vector $u\in R^n$ we denote by $I(u)$ the ideal generated by the components of $u\in R^n$, i.\,e. $I(u)=\sum\limits_{k=1}^nu_kR$.
To prove \cref{a3map} we follow the approach of Tulenbaev and construct yet another family of elements $X_{u,v}(a)\in \St(n,\,B)$, $a\in B$ which contains van der Kallen's generators $X(u, v)$ as a subfamily, i.\,e. $X_{u,v}(1) = X(u, v)$.
The key feature of this definition is that the assumption $u \in\Um(n, B)$ is replaced by a weaker condition $a \in I(u)$ equivalent to imposing that $u$ becomes unimodular after localisation by the powers of $a$.
